---
title: "Visual Inference for Graphcial Diagnostic of Linear Mixed Models"
#subtitle: "⚔<br/>with xaringan"
author: "Kaiwen Jin"
institute: "Monash University"
date: "2020-09-18"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r xaringan-panelset, echo=FALSE}
library(xaringanExtra)
xaringanExtra::use_panelset()
```

# Linguistic Case

.panelset[
.panel[.panel-name[Gender]

```{r include=FALSE}
library(lme4)
library(tidyverse)
library(ggbeeswarm)
library(colorspace)
library(ggplot2)
library(nullabor)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE#,
  #cache.path = "cache/",
  #cache = TRUE
)
df <- read_csv('~/Desktop/Masters/Master_Econometrics/S2 2020/ETF5550/master/data/politeness_data.csv') %>% 
  mutate(scenario = fct_reorder(factor(scenario), frequency, function(x) mean(x, na.rm = TRUE)),
         subject = fct_reorder(subject, frequency, function(x) mean(x, na.rm = TRUE)))
df <- na.omit(df)
```

```{r gender,fig.height=4, dev='svg'}
ggplot(df, aes(gender, frequency)) + geom_point() 
```
]

.panel[.panel-name[Subject]

```{r subject, fig.height=4, dev='svg'}
ggplot(df, aes(attitude, frequency, fill = gender)) + 
  geom_boxplot() + 
  facet_grid(~subject) +
  geom_beeswarm(color = "#383838") +
  scale_fill_discrete_qualitative() 
```

]

.panel[.panel-name[Scenario]

```{r scenario, fig.height=4, dev='svg'}
ggplot(df, aes(scenario, frequency)) + geom_boxplot() + geom_beeswarm(aes(color = gender))
```

]
]
---

# Linear Mixed Model
--
.center[$$\boldsymbol{y = X\beta + Zb + e}$$]
--

where 
  - $\mathbf{y}$ is a $\mathbf{N \times 1}$ vector of observations, outcome variable
  - $\mathbf{X}$ is a $\mathbf{N \times p}$ matrix
  - $\boldsymbol\beta$ is a $\mathbf{p \times 1}$ vector of the fixed effect
  - $\mathbf{b}$ is a $\mathbf{N \times q}$ matrix
  - $\mathbf{b}$ is a $\mathbf{q \times 1}$ vector of the random effect
  - $E[\mathbf{y}] = \boldsymbol{X\beta}$ and $V(\mathbf{y}) = \mathbf\Omega = \boldsymbol{Z\Gamma Z^\top + R}$

How can we implement the LME?
- `lmer` function from `lme4` package
- `mmer` function from `sommer` package

---

# Graphical diagnostic on residual analysis

* Types of residuals and corresponding residual diagnostic purpose:

  - Marginal residuals, $\boldsymbol{\hat\xi = y - X\hat \beta}$
      + Linear of the effects fixed 
      + Presence of outlying observations
      + Within-units covariance matrix 
    
  - Conditional residuals, $\boldsymbol{\hat e = y - X\hat \beta - Z\hat b}$
      + Presence of outlying observations 
      + Homoskedasticity of conditional errors 
      + Normality of conditional errors 
      
  - Random effect residuals, $\mathbf{Z\hat b}$
      + Presence of outlying subjects 
      + Normality of the random effects 

---

# Linguistic Case 

--
### Presence of outlying observations
--

.panelset[
.panel[.panel-name[Subject]
```{r setup, include = FALSE}
fit <- lmer(frequency ~ attitude + gender + 
              (1|subject) + (1|scenario),
            data = df)

N <- nrow(df)
subjects <- unique(df$subject)
nsubject <- length(subjects)
scenarios <- unique(df$scenario)
nscenario <- length(scenarios)
y <- df$frequency
X <- model.matrix(~attitude + gender, data = df)
beta <- fixef(fit)
Z1 <- model.matrix(~subject - 1, data = df)
Z2 <- model.matrix(~scenario - 1, data = df)
Z <- cbind(Z1, Z2)
u <- c(ranef(fit)$subject[,1], ranef(fit)$scenario[,1])
q <- length(u)
vc <- as.data.frame(VarCorr(fit))
R <- vc$vcov[vc$grp=="Residual"] * diag(N)
G1 <- vc$vcov[vc$grp=="subject"] * diag(nsubject)
G2 <- vc$vcov[vc$grp=="scenario"] * diag(nscenario)
G <- Matrix::bdiag(G1, G2)
Sigma <- Z %*% G %*% t(Z)  + R
Sigma_inv <- solve(Sigma)
varMargRes <- Sigma - X %*% solve(t(X) %*% solve(Sigma) %*% X) %*% t(X)
P <- Sigma_inv - Sigma_inv %*% X %*% solve(t(X) %*% Sigma_inv %*% X) %*% t(X) %*% Sigma_inv
varCondRes <- R %*% P %*% R
varU <- G - G %*% t(Z) %*% P %*% Z %*% G


fit_df <- df %>% 
  mutate(fitted = as.vector(X %*% beta),
         resm = frequency - fitted, 
         resc = frequency - fitted - as.vector(Z %*% u),
         index = 1:n()) %>% 
  rowwise() %>% 
  mutate(resm_std = resm / sqrt(diag(varMargRes)[index]),
         resc_std = resc / sqrt(diag(varCondRes)[index])) %>% 
  ungroup()

unit_df <- expand_grid(subjects, scenarios) %>% 
  # cannot follow Vi quit well
  mutate(Vi = map2_dbl(subjects, scenarios, ~{
                        ind <- df %>% 
                          mutate(index = 1:n()) %>% 
                          filter(subject==.x,
                                 scenario==.y) %>% 
                          pull(index)
                        mi <- length(ind)
                        # standardised marginal residual
                        Ei <- solve(varMargRes[ind, ind]) %*% fit_df$resm[ind]
                        sqrt(sum((diag(mi) - Ei %*% t(Ei))^2)) / mi
                      }),
         Mi = as.vector(t(u) %*% solve(varU) %*% u)
         )
```

```{r plot2-1, fig.height=4, dev='svg'}
ggplot(fit_df, aes(index, resm_std, 
                   color = subject)) + 
  geom_point() + 
  geom_hline(yintercept = 0) + 
  scale_color_discrete_qualitative()
```
]

.panel[.panel-name[Scenario]

```{r plot2-2, fig.height=4, dev='svg'}
fit_df %>% 
  arrange(scenario) %>% 
  mutate(index = 1:n()) %>% 
  ggplot(aes(index, resm_std, color = scenario)) + 
  geom_point() + 
  geom_hline(yintercept = 0) + 
  scale_color_discrete_qualitative()
```
]
]
---

# Linguistic Case

--
## Normality of conditional errors
--

```{r plot6, fig.height=4.5, dev='svg'}
library(qqplotr)
ggplot(fit_df, aes(sample = resc_std)) + 
  stat_qq_band(bandType = "pointwise", fill = "#8DA0CB", qprobs = c(0.25, 0.75), alpha = 0.4) +
  stat_qq_line(colour = "#8DA0CB") +
  stat_qq_point() +
  xlab("Normal quantiles") +
  ylab("Standard conditional residuals")
```

---
# Lineup Protocol

```{r simulate, include=FALSE}
# Simulate data for null plots from `fit` model
fit.sim  <- simulate(fit, nsim = 19, seed = 1234)
fit.refit <- lapply(fit.sim, refit, object = fit)
fit.simy <- lapply(fit.refit, function(x) getME(x, "y"))
```

```{r data_nullabor, include=FALSE}
fit.sim.y <- do.call("cbind", fit.simy)
fit.sim.y <- reshape2::melt(fit.sim.y)[-1]
names(fit.sim.y) <- c(".n", "y")
fit.sim.y$.n <- as.numeric(str_extract(fit.sim.y$.n, "\\d+"))
fit.sim.y$attitude <- rep(df$attitude, 19)
fit.sim.y$gender <- rep(df$gender, 19)
fit.sim.y$subject <- rep(df$subject, 19)
fit.sim.y$scenario <- rep(df$scenario, 19)
```

```{r data_simulated, include=FALSE}
simdat <- purrr::map_dfr(1:19, function(i){
  df <- fit.sim.y %>% filter(.n == i) 
  m <- lmer(y ~ attitude + gender + (1|subject) + (1|scenario), data = df)
  
  N <- nrow(df)
  subjects <- unique(df$subject)
  nsubject <- length(subjects)
  scenarios <- unique(df$scenario)
  nscenario <- length(scenarios)
  y <- df$y
  X <- model.matrix(~attitude + gender, data = df)
  beta <- fixef(m)
  Z1 <- model.matrix(~subject - 1, data = df)
  Z2 <- model.matrix(~scenario - 1, data = df)
  Z <- cbind(Z1, Z2)
  u <- c(ranef(m)$subject[,1], ranef(m)$scenario[,1])
  q <- length(u)
  vc <- as.data.frame(VarCorr(m))
  R <- vc$vcov[vc$grp=="Residual"] * diag(N)
  G1 <- vc$vcov[vc$grp=="subject"] * diag(nsubject)
  G2 <- vc$vcov[vc$grp=="scenario"] * diag(nscenario)
  G <- Matrix::bdiag(G1, G2)
  Sigma <- Z %*% G %*% t(Z)  + R
  Sigma_inv <- solve(Sigma)
  varMargRes <- Sigma - X %*% solve(t(X) %*% solve(Sigma) %*% X) %*% t(X)
  P <- Sigma_inv - Sigma_inv %*% X %*% solve(t(X) %*% Sigma_inv %*% X) %*% t(X) %*% Sigma_inv
  varCondRes <- R %*% P %*% R
  varU <- G - G %*% t(Z) %*% P %*% Z %*% G
  
  fitted = as.vector(X %*% beta)
  resm = y - fitted
  resc = y - fitted - as.vector(Z %*% u)
  index = 1:nrow(df)
  resm_std = resm / sqrt(diag(varMargRes)[index])
  resc_std = resc / sqrt(diag(varCondRes)[index])
  
  tibble::tibble(df, fitted, resm_std, resc_std, index)
})
```


```{r turedf, include=FALSE}
colnames(fit_df)[5] <- c("y")
```

```{r lineup_2_1, fig.height=6, fig.width=10, dev='svg'}
qplot(x = index, y = resm_std, color = subject, data = fit_df, geom = "point") %+% 
  geom_hline(yintercept = 0) %+% 
  lineup(true = fit_df, samples = simdat, pos = 9)+ 
  facet_wrap( ~ .sample, ncol=5) +
  xlab(NULL) + ylab(NULL) + 
  theme(axis.text.y = element_blank(), axis.text.x = element_blank(),
        axis.ticks.x = element_blank(), axis.ticks.y = element_blank())
```

---

# Lineup Protocol

```{r lineup_6, fig.height=6, fig.width=10, dev='svg'}
ggplot(fit_df, aes(sample = resc_std)) %+% 
  stat_qq_band(bandType = "pointwise", fill = "#8DA0CB", qprobs = c(0.25, 0.75), alpha = 0.4) %+%
  stat_qq_line(colour = "#8DA0CB") %+%
  stat_qq_point() %+% 
  lineup(true = fit_df, samples = simdat, pos = 9) +
  facet_wrap( ~ .sample, ncol=5)+
  xlab(NULL) + ylab(NULL) + 
  theme(axis.text.y = element_blank(), axis.text.x = element_blank(),
        axis.ticks.x = element_blank(), axis.ticks.y = element_blank())
```

