---
title: "Example with linguistic data"
output:
  html_document:
    code_folding: hide
---

Data is based on [this paper](https://arxiv.org/pdf/1308.5499.pdf). Also see [tutorial here](https://web.stanford.edu/class/psych252/section/Mixed_models_tutorial.html).

```{r setup, include = FALSE}
library(lme4)
library(tidyverse)
library(ggbeeswarm)
library(colorspace)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE#,
  #cache.path = "cache/",
  #cache = TRUE
)

df <- read_csv(here::here("data/politeness_data.csv")) %>% 
  mutate(scenario = fct_reorder(factor(scenario), frequency, function(x) mean(x, na.rm = TRUE)),
         subject = fct_reorder(subject, frequency, function(x) mean(x, na.rm = TRUE)))
```

The data contains `r nrow(df)` observations on the voice pitch (or frequency) from `r length(unique(df$subject))` subjects (3 females and 3 males) under `r length(unique(df$scenario))` scenarios with 2 attitudes (informal or polite). There is one missing value for subject `M4` for scenario 6 with polite attitude.


```{r eda-plots}
ggplot(df, aes(subject, frequency, fill = gender)) + 
  geom_boxplot(width = 0.5) +
  geom_beeswarm(color = "#383838") + 
  scale_fill_discrete_qualitative() 

ggplot(df, aes(attitude, frequency,  fill = gender)) + 
  geom_boxplot(width = 0.5) +
  geom_beeswarm(color = "#383838") + 
  facet_grid(. ~ subject) + 
  scale_fill_discrete_qualitative() 

ggplot(df, aes(gender, frequency)) + 
  geom_boxplot(width = 0.5) +
  geom_beeswarm(color = "#383838") + 
  facet_grid(. ~ attitude) + 
  scale_fill_discrete_qualitative() 

ggplot(df, aes(scenario, frequency)) + 
  geom_boxplot() +
  geom_beeswarm(aes(color = gender)) + 
  scale_color_discrete_qualitative()
```

We fit the following model:

$$\begin{array}\texttt{frequency}_{i} &=& \beta_0 + \beta_1\mathbb{I}(\texttt{attitude}_{i}=\texttt{pol}) + \beta_2 \mathbb{I}(\texttt{gender}_{i}=\texttt{M}) +\\ && u_1\mathbb{I}(\texttt{subject}_i=\texttt{F1})+ u_2\mathbb{I}(\texttt{subject}_i=\texttt{F2}) + u_3\mathbb{I}(\texttt{subject}_i=\texttt{F3}) + u_4\mathbb{I}(\texttt{subject}_i=\texttt{M1})+ u_5\mathbb{I}(\texttt{subject}_i=\texttt{M2})+ u_6\mathbb{I}(\texttt{subject}_i=\texttt{M3}) + \\ && u_7\mathbb{I}(\texttt{scenario}_i=\texttt{1}) + u_8\mathbb{I}(\texttt{scenario}_i=\texttt{2}) + u_9\mathbb{I}(\texttt{scenario}_i=\texttt{3}) + u_{10}\mathbb{I}(\texttt{scenario}_i=\texttt{4}) + u_{11}\mathbb{I}(\texttt{scenario}_i=\texttt{5}) + u_{12}\mathbb{I}(\texttt{scenario}_i=\texttt{6}) +\\&& u_{13}\mathbb{I}(\texttt{scenario}_i=\texttt{7}) + e_i\end{array}$$
assuming that $\boldsymbol{u} = (u_1, ..., u_{13})^\top \sim N(\boldsymbol{0}, \mathbf{G})$ and $\boldsymbol{e} = (e_1, ..., e_{84})^\top = \sim N(\boldsymbol{0}, \sigma^2\mathbf{I}_{84})$, where $\mathbf{G}=\begin{bmatrix}\sigma^2_{\texttt{subject}}\mathbf{I}_6 & \mathbf{0}_{6\times 7}\\ \mathbf{0}_{7\times 6} & \sigma^2_{\texttt{scenario}}\mathbf{I}_7\end{bmatrix}.$


```{r model}
fit <- lmer(frequency ~ attitude + gender + 
              (1|subject) + (1|scenario),
            data = df)

N <- nrow(df)
subjects <- unique(df$subject)
nsubject <- length(subjects)
nscenario <- length(unique(df$scenario))
y <- df$frequency
X <- model.matrix(~attitude + gender, data = df)
beta <- fixef(fit)
Z1 <- model.matrix(~subject - 1, data = df)
Z2 <- model.matrix(~scenario - 1, data = df)
Z <- cbind(Z1, Z2)
u <- c(ranef(fit)$subject[,1], ranef(fit)$scenario[,1])
vc <- as.data.frame(VarCorr(fit))
R <- vc$vcov[vc$grp=="Residual"] * diag(N)
G1 <- vc$vcov[vc$grp=="subject"] * diag(nsubject)
G2 <- vc$vcov[vc$grp=="scenario"] * diag(nscenario)
G <- Matrix::bdiag(G1, G2)
Sigma <- Z %*% G %*% t(Z)  + R
varMargRes <- Sigma - X %*% solve(t(X) %*% solve(Sigma) %*% X) %*% t(X)

fit_df <- df %>% 
  mutate(fitted = as.vector(X %*% beta),
         resm = frequency - fitted, 
         index = 1:n()) %>% 
  rowwise() %>% 
  mutate(resm_std = resm / sqrt(diag(varMargRes)[index])) %>% 
  ungroup()

unit_df <- tibble(subjects) %>% 
  mutate(Vi = map_dbl(subjects, ~{
                        ind <- df %>% 
                          mutate(index = 1:n()) %>% 
                          filter(subject==.x) %>% 
                          pull(index)
                        mi <- length(ind)
                        # standardised marginal residual
                        Ei <- solve(varMargRes[ind, ind]) %*% fit_df$resm[ind]
                        sqrt(sum((diag(mi) - Ei %*% t(Ei))^2)) / mi
                      })
         )

```

## Diagnostic Plot 1: Linearity of fixed effects

```{r plot1}
ggplot(fit_df, aes(fitted, resm_std)) + 
  geom_hline(yintercept = 0) +
  geom_point()

ggplot(fit_df, aes(interaction(attitude, gender), resm_std)) + 
  geom_hline(yintercept = 0) +
  geom_boxplot() 
```

# Diagnostic Plot 2: Prescence of outlying observations

```{r plot2}
ggplot(fit_df, aes(index, resm_std, 
                   color = subject)) + 
  geom_point() + 
  geom_hline(yintercept = 0) + 
  scale_color_discrete_qualitative()

fit_df %>% 
  arrange(scenario) %>% 
  mutate(index = 1:n()) %>% 
  ggplot(aes(index, resm_std, color = scenario)) + 
  geom_point() + 
  geom_hline(yintercept = 0) + 
  scale_color_discrete_qualitative()
```

# Diagnostic Plot 3: Within-units covariance matrix

```{r plot3}
ggplot(unit_df, aes(subjects, Vi)) + 
  geom_point() + 
  geom_hline(yintercept = 0)
```


