---
title: "Case for MathAchieve and MathAchSchool"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(nlme)
library(dplyr)
library(ggplot2)
library(tidyr)
library(purrr)
library(ggbeeswarm)
```

Each row in `MathAchieve` represents each of the 7185 students within 160 schools. Each row in `MathAchSchool` represents each of 160 schools.

[http://www.stat.rutgers.edu/home/yhung/Stat586/Mixed%20model/appendix-mixed-models.pdf]

The variables that we are going to use:
- School
- SES
- MathAch
- Sector
- MEANSES

```{r}
data("MathAchieve")
summary(MathAchieve)
```
```{r}
attach(MathAchieve)
mses <- tapply(SES, School, mean)
detach(MathAchieve)
```

```{r}
data("MathAchSchool")
head(MathAchSchool)
```

Since the `MEANSES` is different in `MathAchieve` and `MathAchSchool`, the new mean of SES for students in each school is generated.

```{r}
Bryk <- as.data.frame(MathAchieve[,c("School", "SES", "MathAch")])
Bryk$meanses <- mses[as.character(Bryk$School)]
sector <- MathAchSchool$Sector
names(sector) <- row.names(MathAchSchool)
Bryk$Sector <- sector[as.character(Bryk$School)]
contrasts(Bryk$Sector)
```

```{r}
fit <- lmer(MathAch ~ SES + Sector + (1|School), data = Bryk)

N <- nrow(Bryk) 
school <- unique(as.character(Bryk$School))
nschool <- length(school)

y <-  Bryk$MathAch
X <- model.matrix(~ SES + Sector, data = Bryk)
beta <- fixef(fit)
Z <- model.matrix(~ School - 1, data = Bryk)
u <- ranef(fit)$School[,1]
q <- length(u)
vc <- as.data.frame(VarCorr(fit))
R <- vc$vcov[vc$grp == "Residual"] * diag(N)
G <- vc$vcov[vc$grp == "School"] * diag(q)
Sigma <- Z %*% G %*% t(Z) + R
Sigma_inv <- solve(Sigma)
varMargRes <- Sigma - X %*% solve(t(X) %*% solve(Sigma) %*% X) %*% t(X)
P <- Sigma_inv - Sigma_inv %*% X %*% solve(t(X) %*% Sigma_inv %*% X) %*% t(X) %*% Sigma_inv
varCondRes <- R %*% P %*% R
varU <- G - G %*% t(Z) %*% P %*% Z %*% G
```

```{r}
fit_bryk <- Bryk %>% 
  mutate(fitted = as.vector(X %*% beta),
         resm = MathAch - fitted, 
         resc = MathAch - fitted - as.vector(Z %*% u),
         index = 1:n()) %>% 
  rowwise() %>% 
  mutate(resm_std = resm / sqrt(diag(varMargRes)[index]),
         resc_std = resc / sqrt(diag(varCondRes)[index])) %>% 
  ungroup()
```

Lesaffre and Verbeke comment that when the within-unit covariance structure is adequate, $V_i = ||I_{m_i} - \varepsilon_i \varepsilon^\top||^2$, where $\varepsilon_i = \hat \Omega ^{-1/2} \hat \xi$ should be close to zero. However, Singer consider replacing $\varepsilon_i$ in $V_i$ with the standardised marginal residuals $\hat \xi_i^* = [\hat V(\hat \xi_i)]^{-1/2} \hat \xi_i$, where $V(\hat \xi_i)$ refers to the diagonal block of $\Omega - X(X^\top \Omega^{-1} X)^{-1} X^\top$ associated to the $i$ unit.

```{r}
unit_df <- expand_grid(school) %>% 
  # cannot follow Vi quit well
  mutate(Vi = map_dbl(school, ~{
                        ind <- Bryk %>% 
                          mutate(index = 1:n()) %>% 
                          filter(School == .x) %>% 
                          pull(index)
                        mi <- length(ind)
                        # standardised marginal residual
                        Ei <- solve(expm::sqrtm(varMargRes[ind, ind])) %*% fit_bryk$resm[ind]
                        sqrt(sum((diag(mi) - Ei %*% t(Ei))^2)) / mi
                      }),
         Mi = as.vector(t(u) %*% solve(varU) %*% u)
         )
```

## Diagnostic Plot 1: Linearity of fixed effects

```{r}
ggplot(fit_bryk, aes(fitted, resm_std)) +
  geom_hline(yintercept = 0) +
  geom_point()
```

## Diagnostic Plot 2: Prescence of outlying observations

```{r}
ggplot(fit_bryk, aes(index, resm_std)) + geom_point() + geom_hline(yintercept = 0)
```

## Diagnostic Plot 3: Within-units covariance matrix

```{r}
ggplot(unit_df, aes(1:nrow(unit_df), Vi)) + geom_point()+ xlab("Unit indicies") + ylab("Modified Lesaffre-Verbeke index")
```

## Diagnostic Plot 4: Prescense of outlying observations using conditional residual

```{r}
ggplot(fit_bryk, aes(index, resc_std)) + geom_point() + geom_hline(yintercept = 0)
```

## Diagnostic Plot 5: Homoskedasticity of conditional errors 

```{r}
ggplot(fit_bryk, aes(fitted, resc_std)) + geom_point() + geom_hline(yintercept = 0)
```

## Diagnostic Plot 6: Normality of conditional errors

```{r}
ggplot(fit_bryk, aes(sample = resc_std)) + 
  geom_qq() + geom_qq_line(color = "red")
```

## Diagnostic Plot 7: Presencse of outlying subjects 

```{r}
ggplot(unit_df, aes(1:nrow(unit_df), Mi)) + geom_point() + xlab("Unit index") + ylab("Manhalanobis distance")
```

## Diagnostic Plot 8: Normality of the random effects

```{r}
ggplot(unit_df, aes(sample = Mi)) + 
  geom_qq(distribution = stats::qchisq,
          dparams = list(df = q)) +
  geom_qq_line(color = "red",
               distribution = stats::qchisq,
               dparams = list(df = q))
```


```{r}
ggplot(MathAchieve, aes(Sex, MathAch)) +
  geom_beeswarm(alpha = .2) +
  facet_grid(. ~Minority)
```

```{r}
length(unique(MathAchieve$School))
```
160 schools with the `MEANSES` is the nummeric vector of the mean SES for the school.

```{r}
df <- MathAchieve %>% mutate(Sex = ifelse(Sex == "Male", 0, 1), Minority = ifelse(Minority == "No", 0, 1))
str(df)
```

```{r}
model <- lmer(MathAch ~ Sex + SES + (1|School) + (1|MEANSES), data = df)
summary(model)
```

